LaMachine: A meta-distribution for NLP software

### 2018-09-10 docs/lamachine_extabstract_clarin2018.tex
removed picture
added architecture schema to paper

%%iris: ik heb dit stukje geintegreerd in de  case study... (ik heb de eerste paragraaf toch nog even hier opgenomen)
Installation of LaMachine begins with a single bootstrap command\footnote{See
\url{https://proycon.github.io/LaMachine}}.  It may interactively query the users for their software preferences, e.g.
the flavour of LaMachine and the set of software to install, which is never static but can be customized by the user.
The user may also opt for installing the latest releases, the more experimental development versions of the software, or
specific custom versions (to facilitate scientific reproducibility). The bootstrap procedure detects and installs
the necessary prerequisites automatically and eventually invokes Ansible to perform the bulk of the work.

%
%After a successful build, the user may interact with LaMachine either through the command line, which offers a standard
%shell and enables access to all lower-level tools and programming languages; or through his or her webbrowser, which
%when enabled offers a simple portal page towards all installed web-capable tools. Amongst these services is also a
%Jupyter Notebook environment\footnote{\url{https://jupyter.org/}}, which offers a more graphical scripting environment
%for e.g. Python and R that has gained great popularity in the scientific community.
%
%The current version of LaMachine comes with some simple data sharing facilities only: we provide a
%single shared dataspace between host and VM/container (where applicable), or allow simple upload. Extensive data
%search and management functions are deliberately beyond the scope of LaMachine, and left to more high-level tooling.

@@

\begin{figure}[htb] \begin{center}
\includegraphics[width=160.0mm]{doublescreenshot.jpg}
\end{center}
\caption{Screenshot excerpts of the Portal Web-app inside LaMachine (left) and the Jupyter Lab environment (right)}
\label{fig:arch}
\end{figure}

@@

LaMachine exists since May 2015 and has been used extensively ever since by numerous users, in 2018 version 2 was
released which was a significant rewrite. LaMachine was initially conceived as the primary means of distribution of the
software stack developed at CLST, Radboud University Nijmegen. It therefore includes a lot of our software.  A full list
of included software goes beyond the scope of this overview; we will merely mention some CLARIN-NL/CLARIAH-funded tools:
ucto (a tokeniser), Frog (an NLP suite for Dutch), FoLiA (Format for Linguistic Annotation, with assorted tools), FLAT
(a web-based linguistic Annotation tool), PICCL (an OCR and post-OCR correction pipeline) and CLAM. However, this
project is not limited to one research group and is open to participation by other software providers, especially those
also in CLARIAH and the upcoming CLARIAH PLUS project. We already include some relevant software by the University of
Groningen (such as Alpino, a dependency parser) and VU Amsterdam. Moreover, LaMachine incorporates a large number of
renowned tools by external international parties, offering most notably a mature Python environment with renowned
scientific modules such as scipy, numpy, scikit-learn, matplotlib, nltk, spacy, pytorch, keras, gensim, tensorflow, and many others,
but also R, Java and tools such as Stanford CoreNLP and Kaldi.

@@

LaMachine comes with some simple data sharing facilities that allowed us to access the sensitive complaint data via a single shared dataspace between host and the VM. (Note that extensive data search and management functions are deliberately beyond the scope of LaMachine, and left to more high-level tooling.)

>>>>

%%iris: ik heb dit stukje geintegreerd in de  case study... (ik heb de eerste paragraaf toch nog even hier opgenomen)
Installation of LaMachine begins with a single bootstrap command\footnote{See
\url{https://proycon.github.io/LaMachine}}.  It can interactively query the users for their software preferences
\emph{(stored as the host configuration)}, e.g. the flavour of LaMachine, as well as the set of software to install,
\emph{the installation manifest}. This set is never static but can be customized by the user.  The user may also opt for installing the latest releases, the more
experimental development versions of the software, or specific custom versions (to facilitate scientific
reproducibility). The bootstrap procedure detects and installs the necessary prerequisites automatically and eventually
invokes Ansible to perform the bulk of the work.  Figure~\ref{fig:arch} provides a schematic view.

%After a successful build, the user may interact with LaMachine either through the command line, which offers a standard
%shell and enables access to all lower-level tools and programming languages; or through his or her webbrowser, which
%when enabled offers a simple portal page towards all installed web-capable tools. Amongst these services is also a
%Jupyter Notebook environment\footnote{\url{https://jupyter.org/}}, which offers a more graphical scripting environment
%for e.g. Python and R that has gained great popularity in the scientific community.
%
%The current version of LaMachine comes with some simple data sharing facilities only: we provide a
%single shared dataspace between host and VM/container (where applicable), or allow simple upload. Extensive data
%search and management functions are deliberately beyond the scope of LaMachine, and left to more high-level tooling.

\begin{figure}[htb] \begin{center}
\includegraphics[width=135.0mm]{architecture.png}
\end{center}
\caption{\footnotesize{A schematic representation of the LaMachine architecture}}
\label{fig:arch}
\end{figure}

@@

@@

LaMachine exists since May 2015 and has been used extensively ever since by numerous users, in 2018 version 2 was
released which was a significant rewrite. LaMachine was initially conceived as the primary means of distribution of the
software stack developed at CLST, Radboud University Nijmegen. It therefore includes a lot of our software.  A full list
of included software goes beyond the scope of this overview; we will merely mention some CLARIN-NL/CLARIAH-funded tools:
ucto (a tokeniser), Frog (an NLP suite for Dutch), FoLiA (Format for Linguistic Annotation, with assorted tools), FLAT
(a web-based linguistic Annotation tool), PICCL (an OCR and post-OCR correction pipeline) and CLAM. However, this
project is not limited to one research group and is open to participation by other software providers, especially those
also in CLARIAH and the upcoming CLARIAH PLUS project. We already include some relevant software by other CLARIAH
partners. Moreover, LaMachine incorporates a large number of renowned tools by external international parties, offering
most notably a mature Python environment with renowned scientific modules such as scipy, numpy, scikit-learn,
matplotlib, nltk, spacy, pytorch, keras, gensim, tensorflow, and many others, but also R, Java and tools such as
Stanford CoreNLP and Kaldi.

@@

LaMachine comes with some simple data sharing facilities that allowed us to access the sensitive complaint data via a
single shared dataspace between host and the VM. Extensive data search and management functions are deliberately beyond the scope of LaMachine, and left to more high-level tooling.
